---
title: "512 Project Code"
output:
  word_document:
    fig_height: 5
    fig_width: 8
date: ""
author: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
options(show.signif.stars = FALSE)

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(car)
library(effects)
library(readxl)
library(mosaic)
library(catstats2)
library(ggResidpanel)
library(easyalluvial)
library(emmeans)
library(patchwork)
library(plotly)
library(dplyr)
theme_set(theme_bw()) #Prevents need for + theme_bw() in ggplots
```

Loading data

```{r}
df <- read.csv("PS_2023.10.28_11.10.58.csv")
```

Removing planets that have a controversial flag (whether the confirmation status of a planet has been questioned in the published literature)

```{r}
filtered_df <- df %>% filter(pl_controv_flag != 1)
```

Remove duplicates based on lowest average distance error and keep the most recent publication date in case of ties

```{r}
filtered_df$PublicationDate <- as.Date(paste0(filtered_df$pl_pubdate, "-01"))
filtered_df <- filtered_df %>%
  mutate(AvgDistError = (sy_disterr1 + sy_disterr1) / 2) %>%
  arrange(pl_name, AvgDistError, PublicationDate) %>%
  group_by(pl_name) %>%
  filter(AvgDistError == min(AvgDistError)) %>%
  slice_tail(n = 1) %>%
  ungroup()
filtered_df <- filtered_df %>%
  arrange(pl_name) %>%
  group_by(hostname) %>%
  slice_head(n=1) %>%
  ungroup()
```

Only keeping the columns that we are interested in

```{r}
compact_df <- filtered_df %>% select(pl_name,hostname,discoverymethod,pl_orbper,pl_orbsmax,pl_rade,pl_bmasse,pl_eqt,pl_orbincl,st_teff,st_rad,st_mass,sy_dist,sy_gaiamag) 
```

Renaming columns to more clear names

```{r}
compact_df <- compact_df %>% dplyr::rename(OrbitalPeriod = pl_orbper,SemiMajorAxis = pl_orbsmax, Radius = pl_rade, Mass = pl_bmasse, Temperature = pl_eqt, OrbitInclination = pl_orbincl,StarTemperature = st_teff, StarRadius = st_rad, StarMass = st_mass, Distance = sy_dist,Magnitude = sy_gaiamag)
```

Visualizing missing data

```{r}
#compact_df <- compact_df %>% filter(StarSpecType!="")
library(mi)
mdf <- missing_data.frame(as.data.frame(compact_df))
image(mdf)
```

Based on the missing data visualization and for the sake of simplicity of the analysis, I will only keep the following variables: planet name(for identification only), discovery method, orbital period, star temperature, star mass, magnitude and distance.

```{r}
final_df <- compact_df %>% select(pl_name,discoverymethod,OrbitalPeriod,StarMass,Distance,Magnitude,hostname)
final_df <- final_df %>% drop_na(pl_name,discoverymethod,OrbitalPeriod,StarMass,Distance,Magnitude,hostname)
mdf <- missing_data.frame(as.data.frame(final_df))
image(mdf)
```
Converting discovery method into factors

```{r}
final_df <- final_df %>% mutate(discoverymethod = factor(discoverymethod))
summary(final_df)
```

```{r}
tally(~discoverymethod, data = final_df)
```
```{r}
final_df <- final_df %>% mutate(discoverymethod = forcats::fct_collapse(discoverymethod,"Timing Variations"= c("Eclipse Timing Variations", "Pulsation Timing Variations", "Transit Timing Variations"),Other = c("Astrometry", "Orbital Brightness Modulation","Imaging")))
tally(~discoverymethod, data = final_df)
summary(final_df)
```

```{r}
final_df <- final_df %>%
 mutate(log_Distance = log(Distance), log_OrbitalPeriod = log(OrbitalPeriod), log_StarMass = log(StarMass))
enhanced_stripchart(data=final_df,log_Distance~discoverymethod)
```
```{r}
library(GGally)
final_df %>% select(discoverymethod,log_OrbitalPeriod,log_StarMass,Magnitude,log_Distance) %>% ggpairs()
```
Remove the outlier in the orbital period since it has very large error

```{r}
final_df <- final_df %>% filter(OrbitalPeriod != 402000000)
final_df %>% select(discoverymethod,log_OrbitalPeriod,log_StarMass,Magnitude,log_Distance) %>% ggpairs()
```
```{r}
summary(final_df)
```


Initial model

```{r}
library(lme4)
initial_model <- lm(log_Distance~(discoverymethod+log_OrbitalPeriod+log_StarMass+Magnitude),data=final_df)
summary(initial_model)
Anova(initial_model)
```
Diagnostic and effects plots

```{r}
resid_panel(initial_model, "R", alpha = 0.3)
```

```{r}
plot(allEffects(initial_model, residuals = T), grid = T)
```
VIF Calculation

```{r}
vif(initial_model)
```

Four-way interaction model
```{r}
model1 <- lm(log_Distance~(discoverymethod*log_OrbitalPeriod*log_StarMass*Magnitude),data=final_df)
summary(model1)
Anova(model1,test.statistic = "F")
```


Backward Step-wise Testing Process
```{r}
model2 <- lm(log_Distance~(discoverymethod*log_OrbitalPeriod*log_StarMass*Magnitude)-discoverymethod:log_OrbitalPeriod:log_StarMass:Magnitude,data=final_df)
Anova(model2)
model3 <- lm(log_Distance~(discoverymethod*log_OrbitalPeriod*log_StarMass*Magnitude)-discoverymethod:log_OrbitalPeriod:log_StarMass:Magnitude-discoverymethod:log_StarMass:Magnitude,data=final_df)
Anova(model3)
model4 <- lm(log_Distance~(discoverymethod*log_OrbitalPeriod*log_StarMass*Magnitude)-discoverymethod:log_OrbitalPeriod:log_StarMass:Magnitude-discoverymethod:log_StarMass:Magnitude-discoverymethod:log_OrbitalPeriod:log_StarMass,data=final_df)
Anova(model4)
model5 <- lm(log_Distance~(discoverymethod*log_OrbitalPeriod*log_StarMass*Magnitude)-discoverymethod:log_OrbitalPeriod:log_StarMass:Magnitude-discoverymethod:log_StarMass:Magnitude-discoverymethod:log_OrbitalPeriod:log_StarMass-discoverymethod:log_OrbitalPeriod:Magnitude,data=final_df)
Anova(model5)
summary(model5)
```


Diagnostics and effects for final model
```{r}
resid_panel(model5, "R", alpha = 0.3)
plot(allEffects(model5), grid = T,multiline=T,ci.style="bands")
#plot(allEffects(model5), grid = T,multiline=T,ci.style="bands",selection=1)
#plot(allEffects(model5), grid = T,multiline=T,ci.style="bands",selection=2)
#plot(allEffects(model5), grid = T,multiline=T,ci.style="bands",selection=3)
#plot(allEffects(model5), grid = T,multiline=T,ci.style="bands",selection=4)
```

Pairwise comparison between the discovery methods in original additive model
```{r}
library(emmeans)
res1 <- emmeans(initial_model,pairwise ~ discoverymethod,adjust="tukey")
res1
plot(res1, comparison=T)
multcomp::cld(res1,alpha=0.05, Letters=LETTERS)
model5.emms <- emmeans(initial_model,"discoverymethod")
model5.emms
contrast(model5.emms,list(gamma1=c(0,0,-1,1))) %>% confint() %>% as.data.frame()
contrast(model5.emms,list(gamma1=c(-1,0,0,1))) %>% confint() %>% as.data.frame()
contrast(model5.emms,list(gamma1=c(-1,1,0,0))) %>% confint() %>% as.data.frame()
contrast(model5.emms,list(gamma1=c(-1,0,0,1))) %>% confint() %>% as.data.frame()
```

